---
title: "Writing_functions"
author: "Amanda Howarth"
date: "11/7/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(rvest)

#code below just makes nice figures# 
knitr::opts_chunk$set(
    echo = TRUE, 
    warning = FALSE,
    fig.width = 8,
    fig.height = 6, 
    out.width = "90%"
)

options(
  ggplot2.continuous.color = "viridis", 
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

Writing functions!

Here's z scores 
```{r}
x = rnorm(n = 30, mean = 4, sd = 2.3)

x_again = rnorm(n = 30, mean =6, sd = 0.3)

y = rnorm(n = 30, mean = 4, sd = 2.3)


(x - mean(x)) / sd(x)
(x_again - mean(x_again)) / sd(x_again)

# we would like to automate this process with functions so u don't have to keep writing over and over 


```
Now here's a fucntion - u named it as an object (z_score)
updated to add conditional statment - if X is not numeric (Character, etc.) and if 
it is wrong it will output X should be numeric 
```{r}
z_score = function(x) {
  
  if(!is.numeric(x)) {
    stop("X should be numeric")
  } else if (length(x<3)) {
    stop("x should be longer than 3")
  }
  
  (x - mean(x)) / sd(x)
  
  
}
```
Try out the function 
you are making x as x_again this time and it ouputs (x - mean(x)) / sd(x) of x_again
using error = True lets u knit even with errors 
```{r, error = TRUE}
z_score(x = x_again)

```
Examples of how the conditional statement works 
```{r, error = TRUE}
z_score(x= 3) 
# NA
z_score(x= "my name is jeff")
```

## multiple outputs
here u want the mean & standard deviation of the x factor 
put it in a tibble to make a dataframe 
```{r}
mean_and_sd = function(input_x) {
  
  if(!is.numeric(input_x)) {
    stop("X should be numeric")
  } else if (length(input_x<3)) {
    stop("x should be longer than 3")
  }
  tibble(
  mean_input = mean(input_x),
  sd_input = sd(input_x)
  )
    
}

#produces mean and standard deviation in a dataframe for whatever input the user out in
# could use a list instead (lists allow u to combine things and output many things)

```
testing above function 
```{r, error = TRUE}
mean_and_sd(input_x = y)

```

